---
title: "Meta analysis on cancer and beverage intake"
author: "Izar de Villasante, Brainvitge"
date: '`r format(Sys.Date(),"%e de %B, %Y")`'
#runtime: shiny
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: 4
    toc_float: yes
  word_document:
    toc: yes
    toc_depth: '4'
  pdf_document:
    keep_tex: yes
    toc: yes
    toc_depth: 4
params:
  
  
  
  
  main_folder: !r getwd()
  #/home/izar/notebooks/
  
  men_color: "aquamarine"  #hcl(h=60,c=100,l=65) 
  women_color: "indianred1"      #hcl(h=15,c=100,l=65) #"#ED813E" # 
  results: ./Results
  Datasets: ./Datasets
  mercedes: "EPIC.PANACEA_log2_factors" 
  excel: "Table1.4_type_bev.xlsx"
  variables: [vol]
    
header-includes: 
  - \usepackage{bbm}
  - \usepackage[spanish]{babel}
---
\newline
\newline


```{r setup, include=FALSE}
require(knitr)
#rmarkdown::render("panacea.Rmd", params = "ask")
opts_knit$set(output.dir = getwd())
options(knitr.duplicate.label = "allow")



# include this code chunk as-is to set options
opts_chunk$set(comment = NA, prompt = TRUE, tidy = FALSE, 
               fig.width = 20, fig.height = 20,echo = TRUE, 
               message = FALSE, warning = FALSE, cache=FALSE,error=TRUE, quiet = TRUE, include=TRUE)
Sys.setlocale("LC_TIME", "C")
```

```{r paquetes, include=FALSE}
if(!(require(readxl))) install.packages("readxl")
if(!(require(devtools))) install.packages("devtools")
if(!(require(lmerTest))) install.packages("lmerTest")
if(!(require(ggplot2))) install.packages("ggplot2")
if(!(require(ggpubr)))install.packages("ggpubr")
if(!(require(grid)))install.packages("grid")

if(!(require(ggrepel)))install.packages("ggrepel")
if(!(require(forcats)))install.packages("forcats")
if(!(require(dplyr)))install.packages("dplyr")
if(!(require(gridExtra)))install.packages("gridExtra")


if(!(require(meta)))install.packages("meta")
if(!(require(metafor)))install.packages("metafor")
if(!(require(dmetar)))devtools::install_github("MathiasHarrer/dmetar") 
if(!require("tidyverse")) install.packages("tidyverse")
if(!require("fs")) install.packages("fs")
if(!require("readxl")) install.packages("readxl")


if(!(require(printr))) {
  install.packages(
    'printr',
    type = 'source',
    repos = c('http://yihui.name/xran', 'http://cran.rstudio.com')
  )
}
```


# 0.Load data:

The starting point is an excel with different sheets for every type of cancer:


```{r load data}
library(readxl)
library("tidyverse")
#Table1.cohort <- read_excel("C:/Users/izar/Downloads/Table 1. General study characteristics of included studies(m_a).xlsx", sheet = "Cohort")
#Table1.CC <- read_excel("C:/Users/izar/Downloads/Table 1. General study characteristics of included studies(m_a).xlsx", sheet = "Case-control")
#Table1.All <- read_excel("Table1_all.xlsx", sheet = "All studies", skip=14)

path <- params$excel

path %>%
  excel_sheets()

datasets <- path %>%
        excel_sheets() %>%
        set_names() %>%
       map(read_excel,
           path = path)


        
#str(datasets)
```
In this case cancer types are from sheet 6 to 11:
```{r,cache=FALSE}
cancer_data <- datasets[c(7:14)]
names(cancer_data)

```

All our datasets contain the same columns:
```{r}
names(cancer_data[[1]])
```


# Meta-Analysis:

In the following steps the different cancer sites will be analyzed discriminating by beverage.
Before starting some considerations should be taken.

For binary outcome data as having/not having a disease (cancer), it is common to use the **Mantel-Haenszel method** *(Mantel and Haenszel 1959; Robins, Greenland, and Breslow 1986)*, to pool results. The formula for this method looks like this:

\begin{equation}
\hat\theta_{MH} = \frac{\sum\limits_{k=1}^K w_{k} \hat\theta_k}{\sum\limits_{k=1}^K w_{k}}
\end{equation}

In order to use Mantel-Haenszel method the following are needed:

* A. **Event rate data**
  - Number of events in the experimental treatment arm
  - Number of participants in the experimental treatment arm
  - Number of events in the control arm
  - Number of participants in the control arm
     
* B. **The so-called person-time** 
  It has to be collected or calculated by hand. What is basically needed to calculate person-time data is:
  - The number of events
  - The timeframe during which they occurred.

Unfortunately, since pre-calculated effect sizes are provided together with Intervals of Confidence the metagen package will be used. In this case, **Mantel-Haenszel method** is not applicable. Instead the **Inverse variance method** is used  $1/\hat\sigma^2_k$ .
The $\sigma^2$ or standard error can be obtained from pre-calculated effect sizes and their interval of confidence. The effect sizes should be transformed into a natural scale using the log-transform before they are pooled as follows:
```{r}
for (i in 1:length(cancer_data)){
  cancer_data[[i]]$RR<- log(as.numeric(cancer_data[[i]]$Value)+0.0001)
  cancer_data[[i]]$lower <- log(cancer_data[[i]]$CI_lower+0.0001)
  cancer_data[[i]]$upper <- log(cancer_data[[i]]$CI_upper+0.0001)
  cancer_data[[i]]$seTE <- (cancer_data[[i]]$upper - cancer_data[[i]]$lower)/3.92

}

saveRDS(cancer_data,"cancer_data.rds")
```

Now the seTE values can be used to calculate the combined effect size using the **inverse variance method** $1/\hat\sigma^2_k$ previously mentioned.

* **Random effects:** Since the populations in the studies are different, a random effect meta-analysis will be conducted. 

* **The Knapp-Hartung(-Sidik-Jonkman) adjustment will be used:** Hartung and Knapp (2001a,b) proposed an alternative method for random effects meta-analysis based on a refined variance estimator for the treatment estimate. Simulation studies (Hartung and Knapp, 2001a,b; IntHout et al., 2014; Langen et al., 2018) show improved coverage probabilities compared to the classic random effects method.

Let's get into it:



```{r,results='asis', quiet = TRUE,include=TRUE}
#lapply(cancer_data, function(x) {
#  knitr::knit_child(
#    'template.Rmd', envir = environment(), quiet = TRUE)
#})
cancer_n <-1
#beverage_n <-1
for (x in cancer_data){
  cat(knit_child(text = readLines('template.Rmd'),envir=globalenv(), quiet = TRUE),echo=FALSE)
  cancer_n <-cancer_n+1
  }
```


# Summaries


```{r}
summaries
```

